\chapter{Macroeconomics}


\section{Introduction to Macroeconomics}
\begin{itemize}
    \item Definition and scope of macroeconomics.
    \item Microeconomics vs. macroeconomics.
    \item Objectives of macroeconomic policy:
    \begin{itemize}
        \item Economic growth.
        \item Price stability.
        \item Full employment.
        \item Balance of payments stability.
    \end{itemize}
\end{itemize}

\section{National Income Accounting}
\begin{itemize}
    \item Measurement of national income:
    \begin{itemize}
        \item Gross Domestic Product (GDP).
        \item Gross National Product (GNP).
        \item Net National Product (NNP).
    \end{itemize}
    \item Methods of calculating GDP:
    \begin{itemize}
        \item Production approach.
        \item Expenditure approach.
        \item Income approach.
    \end{itemize}
    \item Nominal vs. real GDP.
    \item GDP deflator and price indices.
\end{itemize}

\section{Aggregate Demand and Aggregate Supply}
\begin{itemize}
    \item Components of aggregate demand:
    \begin{itemize}
        \item Consumption, investment, government spending, net exports.
    \end{itemize}
    \item Aggregate supply and its determinants.
    \item Short-run vs. long-run aggregate supply.
    \item Equilibrium in the AD-AS model.
\end{itemize}

\section{Economic Growth and Development}
\begin{itemize}
    \item Factors affecting economic growth:
    \begin{itemize}
        \item Human capital, physical capital, technology, institutions.
    \end{itemize}
    \item Theories of economic growth:
    \begin{itemize}
        \item Classical theory.
        \item Solow growth model.
        \item Endogenous growth theory.
    \end{itemize}
    \item Economic development vs. economic growth.
    \item Sustainable development.
\end{itemize}

\section{Unemployment and Inflation}
\begin{itemize}
    \item Types of unemployment:
    \begin{itemize}
        \item Frictional, structural, cyclical, seasonal.
    \end{itemize}
    \item Measurement of unemployment: unemployment rate.
    \item Types of inflation:
    \begin{itemize}
        \item Demand-pull inflation.
        \item Cost-push inflation.
    \end{itemize}
    \item Measurement of inflation: Consumer Price Index (CPI), Producer Price Index (PPI).
    \item Phillips curve: relationship between inflation and unemployment.
\end{itemize}

\section{Money and Banking}
\begin{itemize}
    \item Functions of money.
    \item Money supply and its measurement.
    \item Role of central banks:
    \begin{itemize}
        \item Monetary policy tools: open market operations, reserve requirements, discount rate.
    \end{itemize}
    \item Commercial banking system.
    \item Creation of credit and the money multiplier.
\end{itemize}

\section{Fiscal Policy}
\begin{itemize}
    \item Definition and objectives of fiscal policy.
    \item Tools of fiscal policy:
    \begin{itemize}
        \item Government expenditure.
        \item Taxation.
    \end{itemize}
    \item Budget deficits and public debt.
    \item Fiscal policy and its impact on aggregate demand.
    \item Crowding-out effect.
\end{itemize}

\section{Monetary Policy}
\begin{itemize}
    \item Definition and objectives of monetary policy.
    \item Types of monetary policy:
    \begin{itemize}
        \item Expansionary and contractionary policies.
    \end{itemize}
    \item Monetary policy instruments:
    \begin{itemize}
        \item Interest rates, open market operations, quantitative easing.
    \end{itemize}
    \item Inflation targeting and central bank independence.
    \item Monetary policy transmission mechanism.
\end{itemize}

\section{International Economics and Open Economy Macroeconomics}
\begin{itemize}
    \item Balance of payments: current account, capital account.
    \item Exchange rate systems:
    \begin{itemize}
        \item Fixed, floating, and managed exchange rates.
    \end{itemize}
    \item Purchasing power parity and interest rate parity.
    \item International trade theories:
    \begin{itemize}
        \item Comparative advantage.
        \item Heckscher-Ohlin theory.
    \end{itemize}
    \item Globalization and its impact on economies.
\end{itemize}

\section{Macroeconomic Schools of Thought}
\begin{itemize}
    \item Classical economics.
    \item Keynesian economics.
    \item Monetarist school.
    \item New classical and real business cycle theory.
    \item New Keynesian economics.
\end{itemize}

\section{Macroeconomic Issues and Challenges}
\begin{itemize}
    \item Economic cycles and stabilization policies.
    \item Income inequality and poverty.
    \item Global financial crises.
    \item Climate change and its macroeconomic impact.
    \item Digital currencies and the future of money.
\end{itemize}