\chapter{Contract Law}

\section{Introduction to Contract Law}
\begin{itemize}
    \item Definition and essentials of a valid contract.
    \item Types of contracts:
    \begin{itemize}
        \item Written and oral contracts.
        \item Express and implied contracts.
        \item Bilateral and unilateral contracts.
    \end{itemize}
\end{itemize}

\section{Formation of a Contract}
\begin{itemize}
    \item Offer and acceptance.
    \item Intention to create legal relations.
    \item Consideration:
    \begin{itemize}
        \item Rules of consideration.
        \item Adequacy and sufficiency of consideration.
    \end{itemize}
    \item Capacity to contract:
    \begin{itemize}
        \item Minors, mental incapacity, intoxication.
    \end{itemize}
    \item Legality of purpose.
\end{itemize}

\section{Terms of a Contract}
\begin{itemize}
    \item Express and implied terms.
    \item Conditions, warranties, and innominate terms.
    \item Exclusion clauses and limitation clauses.
    \item Rules of interpretation.
\end{itemize}

\section{Performance and Discharge of Contracts}
\begin{itemize}
    \item Performance of a contract.
    \item Discharge by agreement or consent.
    \item Discharge by frustration.
    \item Discharge by breach of contract.
\end{itemize}

\section{Breach of Contract and Remedies}
\begin{itemize}
    \item Types of breaches:
    \begin{itemize}
        \item Actual breach.
        \item Anticipatory breach.
    \end{itemize}
    \item Remedies for breach of contract:
    \begin{itemize}
        \item Damages: compensatory, punitive, nominal, and liquidated.
        \item Specific performance.
        \item Injunctions.
        \item Rescission.
    \end{itemize}
\end{itemize}

\section{Vitiating Factors in a Contract}
\begin{itemize}
    \item Misrepresentation.
    \item Mistake.
    \item Duress and undue influence.
    \item Illegality and public policy.
\end{itemize}

\section{Special Types of Contracts}
\begin{itemize}
    \item Contracts of agency.
    \item Contracts of guarantee and indemnity.
    \item Contracts of bailment and pledge.
    \item Electronic contracts and digital agreements.
\end{itemize}

\section{Consumer Protection in Contract Law}
\begin{itemize}
    \item Unfair contract terms.
    \item Consumer rights and remedies.
    \item Sale of goods and services regulations.
    \item Distance selling and e-commerce laws.
\end{itemize}

\section{Case Law and Precedents}
\begin{itemize}
    \item Landmark cases in contract law.
    \item Application of case law to real-world scenarios.
\end{itemize}