\chapter{Patent Law and Policy}

\section{Introduction to Patent Law}
\begin{itemize}
    \item Definition and purpose of patents.
    \item History and evolution of patent law.
    \item International patent systems and harmonization.
\end{itemize}

\section{Patentability Requirements}
\begin{itemize}
    \item Novelty.
    \item Inventive step (non-obviousness).
    \item Industrial applicability (utility).
    \item Subject matter eligibility.
\end{itemize}

\section{Patent Application Process}
\begin{itemize}
    \item Preparing a patent application.
    \item Filing procedures:
    \begin{itemize}
        \item National applications.
        \item International applications (PCT system).
    \end{itemize}
    \item Examination and prosecution.
    \item Patent grant and registration.
\end{itemize}

\section{Patent Specifications and Claims}
\begin{itemize}
    \item Structure of a patent document:
    \begin{itemize}
        \item Title, abstract, and drawings.
        \item Detailed description of the invention.
    \end{itemize}
    \item Drafting patent claims:
    \begin{itemize}
        \item Independent and dependent claims.
        \item Functional and means-plus-function claims.
    \end{itemize}
\end{itemize}

\section{Patent Rights and Enforcement}
\begin{itemize}
    \item Exclusive rights granted by patents.
    \item Term of patent protection.
    \item Infringement of patent rights:
    \begin{itemize}
        \item Direct, indirect, and contributory infringement.
    \end{itemize}
    \item Remedies for patent infringement:
    \begin{itemize}
        \item Injunctions.
        \item Damages and accounting for profits.
    \end{itemize}
\end{itemize}

\section{Patent Policy and Economics}
\begin{itemize}
    \item Role of patents in promoting innovation.
    \item Balancing incentives for inventors and public access.
    \item Patent thickets and anti-commons.
    \item Debate on software patents and business method patents.
\end{itemize}

\section{International Patent Systems and Agreements}
\begin{itemize}
    \item Paris Convention for the Protection of Industrial Property.
    \item Patent Cooperation Treaty (PCT).
    \item TRIPS Agreement (Trade-Related Aspects of Intellectual Property Rights).
    \item Regional patent systems:
    \begin{itemize}
        \item European Patent Office (EPO).
        \item African Regional Intellectual Property Organization (ARIPO).
    \end{itemize}
\end{itemize}

\section{Patent Litigation and Dispute Resolution}
\begin{itemize}
    \item Pre-litigation strategies: licensing and settlement.
    \item Litigation process:
    \begin{itemize}
        \item Jurisdiction and choice of forum.
        \item Patent validity challenges.
    \end{itemize}
    \item Alternative dispute resolution (ADR) mechanisms.
\end{itemize}

\section{Emerging Issues in Patent Law and Policy}
\begin{itemize}
    \item Artificial intelligence and patents.
    \item Biotechnology and pharmaceutical patents.
    \item Patents in clean technology and sustainability.
    \item Open innovation and patent pools.
\end{itemize}

\section{Ethics and Controversies in Patent Law}
\begin{itemize}
    \item Ethical considerations in patenting life forms and genes.
    \item Access to essential medicines and compulsory licensing.
    \item Patent trolls and abusive litigation practices.
    \item Public interest and patent waivers during global crises.
\end{itemize}

\section{Landmark Cases in Patent Law}
\begin{itemize}
    \item Key cases shaping patent jurisprudence.
    \item Interpretation of patent claims and scope.
    \item Cases on patent eligibility and policy impact.
\end{itemize}
