\chapter{Global Business and International Management}

\section{Introduction to Global Business}
\begin{itemize}
    \item Definition and scope of global business.
    \item Drivers of globalization:
    \begin{itemize}
        \item Technological advancements.
        \item Trade liberalization and economic integration.
        \item Growth of multinational corporations (MNCs).
    \end{itemize}
    \item Globalization vs. regionalization.
\end{itemize}

\section{International Trade Theories}
\begin{itemize}
    \item Classical trade theories:
    \begin{itemize}
        \item Absolute Advantage (Adam Smith).
        \item Comparative Advantage (David Ricardo).
    \end{itemize}
    \item Modern trade theories:
    \begin{itemize}
        \item Heckscher-Ohlin Theory.
        \item Product Life Cycle Theory.
        \item Porter’s Diamond Model of National Competitive Advantage.
    \end{itemize}
\end{itemize}

\section{Global Business Environment}
\subsection{Economic Environment}
\begin{itemize}
    \item Differences in economic systems: capitalism, socialism, and mixed economies.
    \item Emerging markets and developing economies.
    \item Role of international organizations: World Trade Organization (WTO), International Monetary Fund (IMF), World Bank.
\end{itemize}

\subsection{Political and Legal Environment}
\begin{itemize}
    \item Political systems: democracy, authoritarianism, and totalitarianism.
    \item Legal systems: common law, civil law, and religious law.
    \item International trade regulations and sanctions.
    \item Intellectual property rights and patents.
\end{itemize}

\subsection{Cultural Environment}
\begin{itemize}
    \item Understanding cultural differences:
    \begin{itemize}
        \item Hofstede’s Cultural Dimensions.
        \item Trompenaars’ Seven Dimensions of Culture.
    \end{itemize}
    \item Cross-cultural communication and negotiation.
    \item Managing cultural diversity in international business.
\end{itemize}

\section{Global Strategic Management}
\subsection{Global Strategy Formulation}
\begin{itemize}
    \item Global, multi-domestic, and transnational strategies.
    \item Competitive advantage in global markets.
    \item Strategic alliances, mergers, and acquisitions.
\end{itemize}

\subsection{Market Entry Strategies}
\begin{itemize}
    \item Exporting and importing.
    \item Licensing and franchising.
    \item Joint ventures and wholly-owned subsidiaries.
    \item Greenfield investments vs. acquisitions.
\end{itemize}

\subsection{Global Supply Chain Management}
\begin{itemize}
    \item Managing global logistics and distribution networks.
    \item Risks and challenges in global supply chains.
    \item Sustainable global supply chain practices.
\end{itemize}

\section{International Financial Management}
\begin{itemize}
    \item Exchange rates and currency markets.
    \item Foreign exchange risk management:
    \begin{itemize}
        \item Transaction, translation, and economic risks.
    \end{itemize}
    \item International financing options:
    \begin{itemize}
        \item Foreign direct investment (FDI).
        \item International capital markets.
    \end{itemize}
    \item Role of multinational financial institutions.
\end{itemize}

\section{Global Marketing and Branding}
\begin{itemize}
    \item Global vs. local marketing strategies.
    \item Adapting products and services for global markets.
    \item Branding in international markets.
    \item Digital marketing in a global context.
\end{itemize}

\section{Global Human Resource Management}
\begin{itemize}
    \item Recruitment and selection in international businesses.
    \item Expatriate management and cultural adjustment.
    \item Training and development for global managers.
    \item International compensation and labor relations.
\end{itemize}

\section{Ethics and Corporate Social Responsibility in Global Business}
\begin{itemize}
    \item Ethical dilemmas in global business.
    \item Corporate social responsibility (CSR) in international contexts.
    \item Sustainable business practices in global operations.
    \item Role of global organizations in promoting ethical standards.
\end{itemize}

\section{Emerging Trends in Global Business}
\begin{itemize}
    \item Impact of digitalization and e-commerce.
    \item Global impact of artificial intelligence and automation.
    \item Role of regional trade blocs and agreements:
    \begin{itemize}
        \item European Union (EU), ASEAN, NAFTA/USMCA, etc.
    \end{itemize}
    \item Challenges in global risk management:
    \begin{itemize}
        \item Geopolitical risks, pandemics, and natural disasters.
    \end{itemize}
\end{itemize}