\chapter{Operations and Supply Chain Management}

\section{Introduction to Operations and Supply Chain Management}
\begin{itemize}
    \item Definition and importance of operations management.
    \item Objectives of operations management: efficiency, effectiveness, and adaptability.
    \item Key concepts in supply chain management (SCM).
\end{itemize}

\section{Operations Management}
\subsection{Operations Strategy}
\begin{itemize}
    \item Alignment with business strategy.
    \item Competitive priorities: cost, quality, delivery, flexibility, and innovation.
    \item Product and process design decisions.
\end{itemize}

\subsection{Process Analysis and Design}
\begin{itemize}
    \item Types of processes: job shop, batch, assembly line, and continuous flow.
    \item Process mapping and flowcharts.
    \item Lean processes and Six Sigma principles.
    \item Capacity planning and bottleneck analysis.
\end{itemize}

\subsection{Production Planning and Control}
\begin{itemize}
    \item Forecasting methods:
    \begin{itemize}
        \item Time series analysis.
        \item Causal models and qualitative forecasting.
    \end{itemize}
    \item Aggregate planning:
    \begin{itemize}
        \item Strategies: chase, level, and hybrid.
    \end{itemize}
    \item Master production scheduling (MPS).
    \item Material requirements planning (MRP).
    \item Scheduling and sequencing methods:
    \begin{itemize}
        \item Gantt charts, critical path method (CPM), and program evaluation review technique (PERT).
    \end{itemize}
\end{itemize}

\subsection{Quality Management}
\begin{itemize}
    \item Total Quality Management (TQM).
    \item Statistical process control (SPC) and control charts.
    \item ISO standards (e.g., ISO 9000).
    \item Quality improvement tools: Pareto charts, fishbone diagrams.
\end{itemize}

\section{Supply Chain Management}
\subsection{Introduction to Supply Chains}
\begin{itemize}
    \item Components of a supply chain: suppliers, manufacturers, distributors, retailers, and customers.
    \item Supply chain flows: material, information, and financial.
\end{itemize}

\subsection{Supply Chain Strategy}
\begin{itemize}
    \item Responsive vs. efficient supply chains.
    \item Bullwhip effect and how to mitigate it.
    \item Outsourcing, offshoring, and reshoring decisions.
\end{itemize}

\subsection{Inventory Management}
\begin{itemize}
    \item Types of inventory: raw materials, work-in-process (WIP), and finished goods.
    \item Inventory models:
    \begin{itemize}
        \item Economic order quantity (EOQ).
        \item Reorder point and safety stock.
        \item Just-In-Time (JIT) and Kanban systems.
    \end{itemize}
    \item ABC analysis and inventory classification.
\end{itemize}

\subsection{Logistics Management}
\begin{itemize}
    \item Transportation modes and their trade-offs.
    \item Warehousing and distribution strategies.
    \item Reverse logistics and returns management.
    \item Global logistics and trade regulations.
\end{itemize}

\section{Emerging Trends in Operations and Supply Chain Management}
\begin{itemize}
    \item Digital supply chains and Industry 4.0.
    \item Artificial intelligence and machine learning in operations.
    \item Blockchain for supply chain transparency.
    \item Sustainability and green supply chain practices.
    \item Risk management in global supply chains.
\end{itemize}

\section{Performance Measurement and Metrics}
\begin{itemize}
    \item Key performance indicators (KPIs) for operations:
    \begin{itemize}
        \item Cycle time, throughput, and utilization.
        \item Overall Equipment Effectiveness (OEE).
    \end{itemize}
    \item KPIs for supply chain management:
    \begin{itemize}
        \item Fill rate, order accuracy, and on-time delivery.
        \item Cash-to-cash cycle time.
        \item Inventory turnover ratio.
    \end{itemize}
\end{itemize}

\section{Decision-Making Tools in Operations and Supply Chain Management}
\begin{itemize}
    \item Linear programming and optimization models.
    \item Decision trees and simulation modeling.
    \item Multi-criteria decision-making (MCDM).
\end{itemize}
