\chapter{Corporate Finance and Strategic Financial Management}

\section{Corporate Finance}
\subsection{Capital Budgeting}
\begin{itemize}
  \item Investment decision-making processes:
  \begin{itemize}
    \item Net Present Value (NPV).
    \item Internal Rate of Return (IRR).
    \item Payback period and discounted payback.
    \item Profitability index. 
  \end{itemize}
  \item Cash flow estimation: incremental cash flows, sunk costs, and opportunity costs.
  \item Scenario and sensitivity analysis in project evaluation.
\end{itemize}

\subsection{Cost of Capital}
\begin{itemize}
  \item Weighted Average Cost of Capital (WACC).
  \item Cost of equity (CAPM, Dividend Discount Model).
  \item Cost of debt and preferred stock.
  \item Marginal cost of capital.
\end{itemize}

\subsection{Financing Decisions}
\begin{itemize}
  \item Sources of capital: equity, debt, and hybrid securities.
  \item Optimal capital structure and leverage:
  \begin{itemize}
    \item Modigliani-Miller theorem (with and without taxes).
    \item Trade-off theory, pecking order theory, and signaling theory.
  \end{itemize}
  \item Dividend policy:
  \begin{itemize}
    \item Dividend irrelevance theory (MM hypothesis).
    \item Practical dividend policies and factors influencing them.
  \end{itemize}
\end{itemize}
        
\subsection{Working Capital Management}
\begin{itemize}
  \item Cash management strategies.
  \item Inventory management models (e.g., EOQ).
  \item Accounts receivable management and credit policies.
  \item Financing working capital: trade credit, lines of credit, commercial paper.
\end{itemize}

\subsection{Mergers, Acquisitions, and Restructuring}
\begin{itemize}
  \item Types of mergers (horizontal, vertical, conglomerate).
  \item Valuation of M\&A targets:
  \begin{itemize}
    \item Comparable company analysis (CCA).
    \item Precedent transaction analysis (PTA).
    \item Discounted cash flow (DCF) analysis.
  \end{itemize} 
  \item Synergies, due diligence, and integration planning.
  \item Divestitures, spin-offs, and carve-outs. 
\end{itemize}

\subsection{Risk Management}
\begin{itemize}
  \item Hedging strategies using derivatives:
  \begin{itemize}
    \item Forward contracts, futures, options, and swaps.
  \end{itemize}
  \item Managing interest rate, currency, and credit risk.
  \item Value at Risk (VaR) and other risk measures.
\end{itemize}

\subsection{Corporate Governance}
\begin{itemize}
  \item Role of the board of directors and executive management.
  \item Agency theory and resolving conflicts of interest.
  \item Corporate social responsibility (CSR) and stakeholder management.
\end{itemize}
    

\section{Strategic Financial Management}
\subsection{Strategic Planning and Financial Strategy}
\begin{itemize}
  \item Aligning financial strategy with corporate objectives.
  \item SWOT analysis and Porter's Five Forces applied to financial planning.
  \item Long-term versus short-term financial strategies.
\end{itemize}

\subsection{Business Valuation}
\begin{itemize}
  \item Methods of valuation:
  \begin{itemize}
    \item DCF analysis (free cash flow to the firm, free cash flow to equity).
    \item Comparable company multiples (EV/EBITDA, P/E).
    \item Asset-based valuations.
  \end{itemize}
  \item Adjustments for market conditions, control premiums, and minority discounts.
\end{itemize}
        
\subsection{Performance Measurement and Evaluation}
\begin{itemize}
  \item Economic Value Added (EVA) and Market Value Added (MVA).
  \item Return on Invested Capital (ROIC) and its strategic implications.
  \item Balanced Scorecard and other performance measurement frameworks.
\end{itemize}

\subsection{Financial Forecasting and Modeling}
\begin{itemize}
  \item Building financial models: assumptions, drivers, and outputs.
  \item Sensitivity analysis and scenario planning.
  \item Monte Carlo simulations for risk assessment.
\end{itemize}

\subsection{Strategic Investment and Financing Decisions}
\begin{itemize}
  \item Evaluating and managing strategic investments in innovation, R\&D, and technology.
  \item Real options analysis (e.g., abandonment, expansion, deferral).
  \item Optimal financing mix for strategic investments.
\end{itemize}

\subsection{International Finance}
\begin{itemize}
  \item Foreign exchange risk management: translation, transaction, and economic risk.
  \item Cross-border financing and multinational capital budgeting.
  \item Country risk analysis and political risk management.
\end{itemize}

\subsection{Behavioral Finance and Decision-Making}
\begin{itemize}
  \item Impact of psychological biases on financial decision-making.
  \item Overconfidence, loss aversion, and herd behavior in corporate finance.
  \item Strategies to mitigate cognitive biases.
\end{itemize}

\section{Integration of Corporate Finance and Strategy}
\begin{itemize}
  \item Financial strategies for competitive advantage.
  \item Linking corporate finance to corporate strategy and operations.
  \item Role of financial leadership in driving organizational success.
  \item Crisis management and financial turnaround strategies.
\end{itemize}
