\chapter{Analog to Digital Converter}
An Analog to Digital Converter (ADC) samples a voltage value. It can be used for signals with AC components or sample reading out of DC components. Sensor has either an internal ADC, providing a digital interface, or an external ADC. For example, temperature and humidity sensors provide a difference in voltage level, thus only a DC component, making a cheaper low sampling rate ADC sufficient. Other sensor like an accelerometer could either static or slow changing thus only seen as DC component. An accelerometer could also be seen as a signal, e.g., with shock or vibrations, where the signal could be converted to the frequency domain. For RF signals the ADC provides the signal in the time domain but then later converted to the frequency domain.


\section{Nyquist–Shannon Sampling Theorem}
Nyquist–Shannon sampling theorem defines the minimal sampling rate required to perfectly reconstruct a continues-time signal from its discrete samples, without losing information.
The theorem only applies to band-limited signals, i.e., no high frequency components above a certain maximum frequency $f_{\text{max}}$. Usually this is achieved by low-pass filters.
\begin{equation}
  f_s \geq 2 \times f_{\text{max}}
\end{equation}
where $f_s$ is the sampling frequency.

% TODO: make a graph showing under sampling nyquist sampling and oversampling.

If there are high frequency components it can appear as lower frequencies in the sampled signal, due to a phenomenon called alias. See Figure~\ref{} for the alias effect.
% TODO: graph

Perfect anti-alias filters, i.e., low pass filters, are impossible in the real world. However, the theorem provides a critical guideline for digital signal processing. 

Some systems are intentionally undersampled then, having a low pass filter significantly higher than Nyquist sampling rate, i.e., $f_{\text{max}}$. Then there needs to be some additional mechanism to distinguish the higher frequencies components. In other words determine which \textit{Nyquist zone} it belongs. There are always infinity number of Nyquist zones, however, there should be some anti-alias filtering to ensure that there are only frequency components in zone 0, i.e., $0$ to $f_s/2$. The N-th Nyquist zone is defined as:
\begin{equation}
  n\frac{f_s}{2} \to (n+1)\frac{f_s}{2}
\end{equation}


\section{Representing Samples in the Frequency Domain}
Fast Fourier Transform (FFT) is used to compute the Discrete Fourier Transform (DFT) of a signal, converting from time domain to the frequency domain. The signal will be split into discrete frequency components called bins, each bins value is defined in \ref{eq:bin}.

Each bin represents the 
\begin{equation}\label{eq:bin}
  X[k] = \sum{n=0}^{N-1} x[n] \cdot e^{-i2\pi kn/N}
\end{equation}

The amplitude is the magnitude of $X[k]$, i.e., $|X[k]|$, the bin frequency $k$, and the phase is 
\begin{equation*}
  \theta[k]=\atan2(\Im(X[k]),\Re(X[k])).
\end{equation*}
Arctan2 will result in a phase between $\pm\pi$.

We could use Eulers formula to derive a formula for imaginary and real part.
\begin{align*}
  e^{-i\theta} &= \cos{\theta} - i\sin{\theta} \\
  e^{-i2\pi kn/N} &= \cos(2\pi kn/N) - i\sin(2\pi kn/N) \\
  X[k] &= \sum_{n=0}^{N-1} x[k](\cos(2\pi kn/N) - i\sin(2\pi kn/N)) \\
  X[k] &= \sum_{n=0}^{N-1} x[k]\cos(2\pi kn/N) - i\sum_{n=0}^{N-1} x[k]\sin(2\pi kn/N)) \\
  \Rightarrow& \Re(X[k]) = \sum_{n=0}^{N-1} X[n]\cdot\cos(2\pi kn/N) \\
  \Rightarrow& \Im(X[k]) = \sum_{n=0}^{N-1} X[n]\cdot\sin(2\pi kn/N) \\
\end{align*}




\subsection{Channeliser}
As the name suggests a channeliser splits a signal into multiple channels to process it separately.
Typically, a channeliser uses a FFT to separate the signal components into multiple channels. It is also possible to use digital filters, e.g., polyphase filter banks, to separate the frequency component.




\section{ADC Specifications} %TODO: more details
\paragraph{Resolution (bits)} Number of bits represents the analog value, e.g., 8-bit, 12-bit, 16-bit.

\paragraph{Sampling Rate (SPS)} How many samples the ADC can take per second, e.g., 1 MSPS, 100 kSPS.

\paragraph{Input Range} Voltage range the ADC can digitize, e.g., $0$–$\SI{5}{\volt}$, $\pm\SI{10}{\volt}$.

\paragraph{Integral Non-Linearity (INL)} Deviation of the actual transfer function from an ideal line.

\paragraph{Differential Non-Linearity (DNL)} Variation in step size between adjacent codes.

\paragraph{Signal-to-Noise Ratio (SNR)} Ratio of signal power to noise power, often expressed in dB.

\paragraph{Effective Number of Bits (ENOB)} Actual resolution considering noise and distortion.

\paragraph{Spurious-Free Dynamic Range (SFDR)} Ratio of the signal to the largest spurious component.


% TODO:
%Power Consumption: Critical for battery-powered or portable applications.
%Latency: Time delay between sampling and data availability.
% Dynamic Range


\section{ADC types}
\subsection{Successive Approximation (SAR) ADC}
Usually has a bit rate of 8-16 bits, with moderate speed of 10 kSPS to 5 MSPS. 

\subsection{Delta-Sigma ($\Delta\Sigma$) ADC}
With a bit rate of 16-24 bits makes it ideal for high-precision application. However, its slow sampling rate of 1 SPS to 100 kSPS, it can only be for low speed aplication.

\subsection{Flash ADC}
Flash ADC are very fast with sampling rate ussaly above 100 MSPS. However, it has a low resulution of 4-8 bits.

\subsection{Pipelined ADC}
With speeds between 10 MSPS and 100 MSPS, and moderate to high resolution of 8-16 bits, it excel in application such as Automotive Radar~\cite{AD9081}\cite{TI-ADC12DJ3200}, LIDAR~\cite{AD9208}\cite{TI-ADC08D1520}, SDR~\cite{AD9361}\cite{LTC2175}\cite{TI-ADC32RF45}, spectrum analyzers~\cite{AD9680}\cite{Keysight-M9703A}, oscilloscopes~\cite{AD9625}\cite{Tektronix-DPO70000SX}. % TODO: see over sources


\section{ADC Noise and Error Sources}
\paragraph{Quantization Noise} Error introduced by discretizing a continuous signal.

\paragraph{Thermal Noise} Random noise due to thermal agitation in components.

\paragraph{Aperture Jitter} Timing uncertainty in the sampling instant, critical for high-frequency signals.

\paragraph{Gain and Offset Errors} Deviations in the ADC’s transfer function.

\paragraph{Clock Jitter} Uncertainty in the sampling clock, affecting high-speed ADCs.


\subsection{DC Bin}
With converting the sampled signal to the frequency domain, there will be a DC component present in the signal. The signal is typically not centred around 0 volts when sampling. The offset will result in a 0 Hz competent, i.e., a DC component. Therefore, a signal will be shown in the first bin or at 0 Hz.
The first bin is derived from equation~\ref{eq:bin} and is defined in equation~\ref{eq:bin0}.

\begin{equation}\label{eq:bin0}
  X[0] = \sum_{n=0}^{N-1} x[n]
\end{equation}



\section{ADC Interfacing}

\paragraph{Reference Voltage} Denoted as $V_{\text{REF}}$, sets the full-scale input range of the ADC.
% TODO: I hold explain full-scale somewhere 

\paragraph{Input Impedance} Must match the source impedance to avoid signal distortion.

\paragraph{Driving the ADC} Use op-amps or buffers to ensure the input signal is stable during conversion.

\paragraph{Multiplexing} Using a single ADC to sample multiple channels via a multiplexer.




\section{Sampling}
\subsection{Real-Time Sampling}
For real-time control and monitoring.

\subsection{IQ Sampling (In-Phase and Quadrature Sampling)}
For RF and wireless communication.

\subsection{Nyquist Sampling}

\subsection{Oversampling}
For high-fidelity and precision measurements.

\subsection{Undersampling (Bandpass Sampling)}
For radar.


\subsection{Random sampling (Non-Uniform Sampling)}
For sensors?


\subsection{Compressed Sampling (Compressive Sampling)}

\subsection{Event-Based Sampling}

\subsection{Block Sampling}

\subsection{Synchronous vs. Asynchronous Sampling}



\section{Advanced Techniques}

\paragraph{Dithering} Adding noise to improve linearity and reduce distortion in high-resolution ADCs.

\paragraph{Decimation} Reducing the sampling rate after oversampling to improve resolution.

\paragraph{Interleaving} Using multiple ADCs in parallel to increase effective sampling rate.

\paragraph{Calibration} Compensating for ADC errors (gain, offset, INL/DNL) in software or hardware.



