\chapter{Linux}


\section{Linux Kernel Overview}
% Kernel Modules, Drivers


\section{Boot Process}

\subsection{Pre Linux Kernel Boot Process}
\subsection{Linux Kernel Boot Process}
\subsection{System Initialization}


\section{File System}
\subsection{Filesystem Hierarchy Standard}
The filesystem hierarchy standards~\cite{fhs_3.0} defines the file system standard from the root directory. For the user specific directory standard XDG Base Directory Specification~\cite{xdg_basedir_spec} is used.

\dirtree{%
  .1 /\DTcomment{The root directory of the file system.}.
.2 bin/\DTcomment{Binaries of commands for all users.}.
.2 boot/\DTcomment{Contains required files for booting.}.
.2 dev/\DTcomment{Device files used by software to communicate with hardware.}.
.2 etc/\DTcomment{Host-specific system configuration.}.
.2 home/\DTcomment{Directories for all users.}.
.3 \textless{}your systems' username\textgreater{}/\DTcomment{Users home directory.}.
.4 .cache/\DTcomment{Stores non-essential, temporary data, e.g., downloaded packages, or logs.}.
.4 .config/\DTcomment{Stores user-specific configuration files.}.
.4 .local/\DTcomment{Stores user-specific data, configurations, and executables.}.
.4 .ssh/\DTcomment{Contains configuration and key files for SSH connections}.
.4 .var/\DTcomment{Used for runtime files for user applications.}.
.2 lib/\DTcomment{Shared libraries and kernel modules.}.
.2 media/\DTcomment{Removable media, e.g., USB drives, CDs.}.
.2 mnt/\DTcomment{Temporary mount points for filesystems.}.
.2 opt/\DTcomment{Third-party or optional software packages.}.
.2 proc/\DTcomment{Virtual filesystem providing runtime system information, e.g., /proc/cpuinfo.}.
.2 root/\DTcomment{The home directory for the root (superuser) account.}.
.2 run/\DTcomment{Stores runtime data, e.g., process IDs, sockets.}.
.2 sbin/\DTcomment{System administration binaries, e.g., fdisk, ifconfig.}.
.2 selinux/\DTcomment{Used by SELinux to store policies, status, and configuration files.}.
.2 srv/\DTcomment{Data for services provided by the system, e.g., web server files.}.
.2 sys/\DTcomment{A virtual filesystem that exposes kernel and hardware information in a structured way.}.
.2 tmp/\DTcomment{Temporary files created by applications.}.
.2 usr/\DTcomment{Contains user programs, libraries, and documentation.}.
.2 var/\DTcomment{Stores variable data like logs (/var/log), caches, and spool files.}.
}




\section{Systemd and its Alternatives}
During the kernel boot process the kernel is hard coded to initiate the first user process, the init system, usually systemd. The init system is responsible for initiate the other user process. It will always have the process id (PID) 1. However, not all processes listed in the system, e.g., using \textit{top} or \textit{ps aux}, are created by PID 1, there are also kernel threads, e.g., kthreadd, ksoftirqd/0, and cpuhp/0. Each kernel thread is listed within brackets [], unlike user space process, and are almost always listed before the user space process.   

Although systemd has become the de facto standard on Linux. There are other init systems, two of the most popular alternatives are OpenRC and runit.


% .service, .device, ....
\subsection{Systemd}
\subsubsection{Design Philosophy}
Modern, all-in-one suite with aggressive integration, aiming to unify system management.
Key Features:


\subsubsection{Key Features}
\begin{itemize}
\item \textbf{Service Management}: Uses .service files, supports parallel startup, socket activation, and dependency-based ordering.
\item \textbf{Scope}: Beyond init, it manages devices (udev), logging (journald), login sessions (logind), network (networkd), timers, and more.
\item \textbf{Complexity}: Highly feature-rich, often criticized for its monolithic design and tight coupling with Linux.
\item \textbf{Adoption}: Default in most major Linux distributions (e.g., Debian, Ubuntu, Fedora, Arch).
\end{itemize}


\subsubsection{Pros}
\begin{itemize}
\item Fast boot times (parallelization).
\item Advanced features (e.g., cgroups, resource control, snapshots).
\item Strong integration with modern Linux ecosystems.
\end{itemize}


\subsubsection{Cons}
\begin{itemize}
\item Complexity and bloat.
\item Controversial for deviating from Unix philosophy (``do one thing well'').
\end{itemize}



\subsection{OpenRC}
\subsubsection{Design Philosophy}
Lightweight, dependency-based init system, focusing on simplicity and compatibility.


\subsubsection{Key Features}
\begin{itemize}
  \item \textbf{Service Management}: Uses shell scripts in /etc/init.d/, with dependency-based startup (defined in /etc/runlevels/).
  \item \textbf{Scope}: Pure init system; relies on external tools for logging (e.g., syslog), device management (e.g., eudev), and networking.
  \item \textbf{Compatibility}: Works with or without systemd (e.g., used in Alpine Linux, Gentoo, Devuan).
  \item \textbf{Performance}: Lightweight, but slower than systemd for parallel startup.
\end{itemize}


\subsubsection{Pros}
\begin{itemize}
  \item Simple, modular, and transparent.
  \item No hard dependencies on specific Linux features.
  \item Easy to debug (shell scripts).
\end{itemize}


\subsubsection{Cons}
\begin{itemize}
  \item Slower boot times (less parallelization).
  \item Fewer built-in features (e.g., no native socket activation).
\end{itemize}



\subsection{Runit}

\subsubsection{Design Philosophy}
Minimalist, Unix-like, focusing on reliability and simplicity.

\subsubsection{Key Features}
\begin{itemize}
\item \textbf{Service Management}: Uses directories (/etc/sv/<service>/) with run scripts. Supervised by runsvdir.
\item \textbf{Scope}: Pure process supervisor; relies on external tools for everything else (e.g., logging, networking).
\item \textbf{Design}: Follows the "do one thing well" principle. Services are isolated and easy to manage.
\item \textbf{Adoption}: Used in Void Linux, Artix Linux, and as an alternative in other distros.
\end{itemize}


\subsubsection{Pros}
\begin{itemize}
\item Extremely lightweight and fast.
\item Simple to configure and debug.
\item No dependencies on Linux-specific features (portable).
\end{itemize}


\subsubsection{Cons}
\begin{itemize}
\item Minimal built-in features (e.g., no native dependency resolution).
\item Requires manual setup for logging, networking, etc
\end{itemize}


\subsection{Creating Systemd Services}

%\subsection{Managing Systemd Services}



\section{Security}
\subsection{SELinux}
\subsection{User Groups and Permissions}
\subsubsection{SUDO}
\subsubsection{File and Folder Permissions}
\subsubsection{/Home}
\subsection{Firewalld}


\section{Display Protocol}
\subsection{X11}
\subsection{Wayland}


\section{Common Linux Widget Tools}
\subsection{GTK}
\subsubsection{GTK3}
\subsubsection{GTK4}

\subsection{QT}


\section{Desktop Environments}
\subsection{Gnome}
\subsection{KDE}
\subsection{Tiling Window Managers}


\section{Useful Terminal Commands}
\subsection{Output Manipulation}
\paragraph{Less}
\paragraph{Cat}
\paragraph{Write $>$}
\paragraph{Reed $<$}
\paragraph{Pipe $|$}
\paragraph{Grep}
\paragraph{Awk}
\paragraph{Sed}
\paragraph{Cut}
\subsection{Service Management}
\paragraph{Enable/Disable}
\paragraph{Start/Stop}
\paragraph{Status}
\subsection{Networking}
\paragraph{Ping}
\paragraph{Ifconfing}
\paragraph{Ip add}
\subsection{Documentation}
\paragraph{Man}
\subsection{Editing Files}
\paragraph{Nano}
\paragraph{Vi}
\paragraph{Vim}
\subsection{Moving In the Files System}
\paragraph{Find}
\paragraph{Mkdir}
\paragraph{Mv}
\paragraph{Rm}
\paragraph{Ls}
\paragraph{Cp}



\section{Yocto Project}
\begin{description}

    \item[Configuration Files]  
    These files define global variables, user settings, and hardware-specific parameters.  
    They instruct the build system which platform to support and what to include in the final image.

    \item[Recipe]  
    A recipe is the most common metadata unit. It specifies where to fetch source code,  
    which patches to apply, dependencies on libraries or other recipes, and how to configure  
    and compile the software. Recipes are stored within layers.

    \item[Layer]  
    A layer groups related recipes. Layers let you organize and reuse metadata across  
    different hardware architectures or customizations. Layers are hierarchical — later  
    layers can override earlier ones. The Yocto Project provides a searchable Layer Index.

    \item[Metadata]  
    Metadata includes recipes, configuration files, and build instructions. It controls  
    what gets built, how, and from where. Metadata may also contain patches or auxiliary  
    files to fix bugs or add features. OpenEmbedded Core is a core set of validated metadata.

    \item[OpenEmbedded Core (OE-Core)]  
    A foundational set of recipes, classes, and files shared across many embedded Linux  
    systems. OE-Core is a quality-assured subset of the original OpenEmbedded repository,  
    continuously tested for reliability.

    \item[Poky]  
    A reference distribution and test configuration. Poky serves three purposes:  
    (1) demonstrating how to customize a Linux distribution, (2) validating the Yocto  
    Project components, and (3) providing an easy starting point for users. Poky is not  
    production-ready but is ideal for experimentation. It builds on top of OE-Core.

    \item[BitBake (Build System)]  
    A task scheduler and execution engine. BitBake parses recipes and configuration data,  
    builds a dependency tree, schedules compilation, and produces a custom Linux image.  
    It works like \texttt{make} but is tailored for cross-compilation. During a build,  
    BitBake first creates a cross-compiler toolchain (an Extensible SDK) for the target platform.

    \item[Packages]  
    The intermediate outputs produced by the build system, which are assembled into the  
    final image.

    \item[Extensible SDK (eSDK)]  
    A customized SDK for application developers. It allows developers to modify libraries  
    or code and immediately integrate those changes back into the image, making them  
    available to other applications.

    \item[Image]  
    A binary file containing a complete Linux distribution (operating system), ready to  
    be written onto a target device.

\end{description}


\section{Real-Time Linux}




\section{Linux Organizational Structure}
\subsection{Linux Foundation}
\subsection{GNU}
\subsection{Red Hat}
\subsection{SUSE}
\subsection{Canonical}


\section{Linux Distributions}
\subsection{Debian}
\subsubsection{Ubuntu}
\subsection{Red Hat}
\subsubsection{Fedora}
\subsubsection{Red Hat Enterprise Linux}
\subsection{SUSE}
\subsubsection{SUSE Enterprise Linux}
\subsubsection{OpenSUSE}
\paragraph{Tumbelweed}
\paragraph{Leap}
\subsubsection{Enthusiasts Operating Systems}
\paragraph{NixOS}
\paragraph{Arch}
\paragraph{GNU Guix}
\paragraph{Gentoo}


\section{Common System Applications}
% For printing, bluetooth, network, sound, 


\section{Linux Contribution Workflow}
\subsection{Compiling the Kernel Yourself}
\subsection{Rust and C}
\subsection{Drivers}
\subsection{Kernel Modules}

